% Chapter 4

\chapter{Challenges, Threats, and Defenses}
\label{ch:chapter4}

While FL improves privacy at a high level, it introduces new vulnerabilities and system challenges \cite{rodriguez2020federated}.

\section{System-Level Challenges}

\subsection{Statistical Heterogeneity (Non-IID Data)}

Data across clients is often highly diverse. For example, medical data from different hospitals may follow completely different distributions. This leads to poor convergence and reduced model performance.

\subsection{System Heterogeneity}

Devices differ in:
\begin{itemize}
    \item Computational power
    \item Network connectivity
    \item Battery constraints
\end{itemize}

This results in delays, dropped updates, and inefficient training.

\subsection{Communication Overhead}

FL requires frequent communication between server and clients, which becomes costly at scale.

\section{Threat Model}

Federated systems introduce a broader attack surface:

\begin{itemize}
    \item \textbf{Insider attackers:} Clients or server with access to internal information
    \item \textbf{Outsider attackers:} Intercept communication
    \item \textbf{Honest-but-curious:} Follow protocol but attempt to extract information
    \item \textbf{Malicious attackers:} Intentionally corrupt the system
\end{itemize}

\section{Integrity Attacks}

\subsection{Data Poisoning}

Attackers manipulate training data to degrade model performance.

\subsection{Model Poisoning}

Attackers directly alter model updates to influence the global model.

\subsection{Backdoor Attacks}

A particularly dangerous attack where:
\begin{itemize}
    \item Model behaves normally on standard inputs
    \item Produces incorrect outputs for specific trigger patterns
\end{itemize}

These attacks are stealthy and difficult to detect.

\section{Privacy Attacks}

Even without access to raw data, attackers can infer sensitive information:

\subsection{Gradient Inversion}

Reconstructs training data from gradients \cite{zhu2019deep, geiping2020inverting}.

\subsection{Membership Inference}

Determines whether a data point was part of training.

\vspace{0.5cm}
These attacks highlight a critical insight:

\begin{quote}
\textit{Sharing gradients is not equivalent to preserving privacy.}
\end{quote}

\section{Defense Mechanisms}

Several defenses exist but none are universally effective:

\begin{itemize}
    \item Robust aggregation (median, trimmed mean)
    \item Differential privacy noise addition
    \item Gradient compression
    \item Anomaly detection for malicious clients
\end{itemize}

\textbf{Key Limitation:} No single defense can protect against all attack types simultaneously.
